---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r echo=TRUE}
# load the lattice library
library(lattice)

# set the url
url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

# set the destination file
destfile = "activity.zip"

# set the method to curl as it's https we are getting
method = "curl"

# download the file
# download.file(url, destfile, method, quiet = FALSE, mode = "w",
#               cacheOK = TRUE,
#               extra = getOption("download.file.extra"))

# unzip it
unzip(destfile)

# read in the CSV
data = read.csv("activity.csv")
```

```{r echo=TRUE}

# aggregate the total steps over days, disregarding NAs
total_steps_per_day = aggregate(steps ~ date , data = data, FUN = sum, na.rm = TRUE)

# aggregate the total steps over intervals, disregarding NAs
total_steps_per_interval = aggregate(steps ~ interval , data = data, FUN = mean, na.rm = TRUE)

```

## What is mean total number of steps taken per day?

```{r echo=TRUE}
total_steps_per_day_mean = mean(no_missing_data$steps)
```

Mean total steps per day: `r total_steps_per_day_mean`

### make a histogram of the number of steps

```{r histogram, fig.width=8, fig.height=6, fig.path='figure/'}
histogram(total_steps_per_day$steps)
```

### the mean and median steps per day

```{r echo=TRUE}

# calculate the mean 
avg_steps_mean = mean(total_steps_per_day$steps)

# calculate the median
avg_steps_median = median(total_steps_per_day$steps)
```

Calculated Mean of average steps per day: `r avg_steps_mean`

Calculated Median of average steps per day: `r avg_steps_median`

## What is the average daily activity pattern?

The time series plot of the 5-minute interval

```{r, fig.width=8, fig.height=6, fig.path='figure/'}
xyplot(ts(total_steps_per_interval$steps))
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=TRUE}

```

`r `

## Imputing missing values

1. The number of missing values in the dataset:

```{r echo=TRUE}
# find the missing values and count them
total_missing_steps = length(which(is.na(data$steps)))
```

The data has: `r total_missing_steps` missing values.

2. Strategy for filling missing values:

```{r echo=TRUE}
# take the 95th percentile of the good data and use that to substitute out the missing data
steps_95th = quantile(no_missing_data$steps, 0.95)

```

3. Creating a new data set with the NA steps filled in:

```{r echo=TRUE}
# get a data set for augmentation
augmented_data = data

# substitude that into the missing data steps
augmented_data$steps[is.na(augmented_data$steps)] <- steps_95th

total_augmented_steps_per_day = aggregate(steps ~ date, data = augmented_data, FUN = sum)

```

4. Histogram with Mean and Median calculations:

```{r augmented_histogram, fig.width=8, fig.height=6, fig.path='figure/'}
histogram(total_augmented_steps_per_day$steps)
```

```{r echo=TRUE}
# mean calculation
augmented_mean = mean(total_augmented_steps_per_day$steps)

# meadian
augmented_median = mean(total_augmented_steps_per_day$steps)
```

Mean for augmented data: `r augmented_mean`

Median for augmented data: `r augmented_median`


## Are there differences in activity patterns between weekdays and weekends?

1. Get a new vector that has a weekend or weekday

```{r echo=TRUE}
wdata = augmented_data

wdata$wend <- as.factor(ifelse(weekdays( as.Date(data$date)) %in% c("Saturday","Sunday"), "Weekend", "Weekday"))

```

2. Plot them:

```{r echo=TRUE, fig.path='figure/'}
xyplot(ts(wdata[wdata$wend == "Weekend", ]$steps))
xyplot(ts(wdata[wdata$wend == "Weekday", ]$steps))
```

